---
title: testing mathjax/latex
author: Robert Hazell
date: '2020-04-01'
slug: testing-mathjax-latex
categories: []
tags: []
description: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, comment=NA)
# change path to load data and saved models
knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/SMU/bike-sharing-dataset/CapitalBikeShare")) 
```


$PredictedOriginal_{t+1} = PredictedDiff_{t+1} + Actual_t$

$R^{2}$


[El-Assi, et al 2017](https://www.researchgate.net/profile/Khandker_Nurul_Habib2/publication/286379583_Effects_of_Built_Environment_and_Weather_on_Bike_Sharing_Demand_A_Station_Level_Analysis_of_Commercial_Bike_Sharing_in_Toronto/links/56684b7c08ae7dc22ad1c992/Effects-of-Built-Environment-and-Weather-on-Bike-Sharing-Demand-A-Station-Level-Analysis-of-Commercial-Bike-Sharing-in-Toronto.pdf) perform a similar types of analysis of Toronto's bike share program, developing models for weekday, weekend, station origin, and station destination ridership.  Adjusted ${R^2}$ values hover around 0.66.

[Wang et al 2012](https://pdfs.semanticscholar.org/8a39/c1df91725ac27ea0c413360c3592b911d17a.pdf) use 19 total socioeconomic variables to develop a regression model for station level, 2011 ridership of the Nice Ride bike share program in Minneapolis.  Notably no weather variables are included.  They achieve an adjusted Adjusted ${R^2}$ of approximately 0.87.    ${R^2}$


```{r}
summary(lm(mpg~disp, data=mtcars))
```

 ${R^2}$
