---
title: Working around LaTeX and Plotly issue in Hugo Minimal
author: Robert Hazell
date: '2020-04-04'
slug: correcting-latex-issues-with-plotly-in-hugo-minimal
categories: []
tags:
  - out-of-the-loop
description: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, comment=NA)
```

I've noticed some wierd quirks about [Hugo Minimal](https://themes.gohugo.io/minimal/), which is the theme used for this website.  There's no indication it supports by default rendering of LaTeX with MathJax.  However, I have no problem rendering LaTeX - see [this post](https://statisticalmusings.netlify.com/post/mean-variance-standardized-data/) of mine for example.

What *is* problematic for me is including a plotly chart in a document that contains LaTeX.  Plotly is nice if you want to include interactive plots in your documents. Here's a scatterplot from the ```mtcars``` dataset.

```{r}
library(plotly)
plot_ly(mtcars,
        x = ~hp,
        y = ~mpg)
```

For some reason if I include a plotly chart, LaTeX typed in the document will not render.  

$c^2 = a^2 + b^2 - 2abcos(C)$

As shown above, you instead see the raw LaTeX.

The workaround that checks-out for me is using a LaTeX online editor like [CodeCogs](https://www.codecogs.com/latex/eqneditor.php).  The website converts math expressions you type to LaTeX.  They automatically produce ```html``` code that can be copied and pasted into your document -- even inline!  So, I can write the law of cosines formula is <a href="https://www.codecogs.com/eqnedit.php?latex=\inline&space;c^2&space;=&space;a^2&space;&plus;&space;b^2&space;-&space;2abcos(C)" target="_blank"><img src="https://latex.codecogs.com/gif.latex?\inline&space;c^2&space;=&space;a^2&space;&plus;&space;b^2&space;-&space;2abcos(C)" title="c^2 = a^2 + b^2 - 2abcos(C)" /></a> with no issue, thanks to the ```html``` produced by CodeCogs.  It looks like this:

```{r eval=FALSE}
<a href="https://www.codecogs.com/eqnedit.php?latex=\inline&space;c^2&space;=&space;a^2&space;&plus;&space;b^2&space;-&space;2abcos(C)" target="_blank"><img src="https://latex.codecogs.com/gif.latex?\inline&space;c^2&space;=&space;a^2&space;&plus;&space;b^2&space;-&space;2abcos(C)" title="c^2 = a^2 + b^2 - 2abcos(C)" /></a>
```

So I don't know why LaTeX code renders natively without modifying Minimal's Javascript, or why it it *doesn't* render when I include plotly charts.  One of those is a happy coincidence.  Thankfully for the other one, you can turn to CodeCogs to sort it out!  