---
title: bankingr
author: ''
date: '2020-04-23'
slug: bankingr
categories: []
tags: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, comment=NA)
```

```{r}
library(reticulate)
use_python("/usr/local/bin/python3", required = T)
```

```{python}
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt
bankingDF = pd.read_csv("/Users/roberthazell/Desktop/SMU/ML1/bank-additional-full.csv")
```

## Business Understanding

The Bank Marketing Data Set from the UCI Machine Learning Repository was selected to be analyzed by our group. The data involves information obtained from direct telemarketing campaigns from a Portuguese banking institution.  The data contains 41,188 observations with 20 attributes and was collected between May 2008 to November 2010.

Our objective is to build a customer conversion model that predicts the likelihood a customer converts after seeing a bank term deposit offer.  A cash investment, which will be held at the bank, is invested at an agreed rate of interest over a fixed period of time.  Characteristics of a customer such as age, education, type of job, and many others will be analyzed to determine if they are factors determining conversion.

A good prediction algorithm would establish a relationship or correlation between specific attributes with the probability of whether a bank customer would subscribe to a bank term deposit.  We will assess the qualities of our models using a confusion matrix.  With the results from our analysis, the attributes that are strongly correlated to a customer subscribing to a bank term deposit would be emphasized in future telemarketing campaigns.

## Data Understanding 

### Data Dictionary

<style type="text/css">
.tg  {border-collapse:collapse;border-spacing:0;border-color:#ccc;}
.tg td{font-family:Arial, sans-serif;font-size:14px;padding:10px 5px;border-style:solid;border-width:0px;overflow:hidden;word-break:normal;border-top-width:1px;border-bottom-width:1px;border-color:#ccc;color:#333;background-color:#fff;}
.tg th{font-family:Arial, sans-serif;font-size:14px;font-weight:normal;padding:10px 5px;border-style:solid;border-width:0px;overflow:hidden;word-break:normal;border-top-width:1px;border-bottom-width:1px;border-color:#ccc;color:#333;background-color:#f0f0f0;}
.tg .tg-fymr{font-weight:bold;border-color:inherit;text-align:left;vertical-align:top}
.tg .tg-0pky{border-color:inherit;text-align:left;vertical-align:top}
</style>
<table class="tg">
  <tr>
    <th class="tg-fymr">Attribute</th>
    <th class="tg-fymr">Data Type</th>
    <th class="tg-fymr">Description</th>
    <th class="tg-fymr">Examples</th>
  </tr>
  <tr>
    <td class="tg-0pky">age</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">The age of the customer in years</td>
    <td class="tg-0pky">26, 45, 50</td>
  </tr>
  <tr>
    <td class="tg-0pky">job</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">The type of job of the customer</td>
    <td class="tg-0pky">admin, blue-collar, entrepreneur, housemaid, <br>management, retired, self-employed, services, <br>student,technician, unemployed, unknown</td>
  </tr>
  <tr>
    <td class="tg-0pky">marital</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">The marital status of the customer</td>
    <td class="tg-0pky">divorced, married, single, unknown</td>
  </tr>
  <tr>
    <td class="tg-0pky">education</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">The education level of the customer</td>
    <td class="tg-0pky">basic.4y, basic.6y, basic.9y, <br>high.school, illiterate, professional.course, <br>university.degree, unknown</td>
  </tr>
  <tr>
    <td class="tg-0pky">default</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">If the customer has their credit in default</td>
    <td class="tg-0pky">yes, no</td>
  </tr>
  <tr>
    <td class="tg-0pky">housing</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">If the customer has a housing loan</td>
    <td class="tg-0pky">yes, no</td>
  </tr>
  <tr>
    <td class="tg-0pky">loan</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">If the customer has a personal loan</td>
    <td class="tg-0pky">yes, no</td>
  </tr>
  <tr>
    <td class="tg-0pky">contact</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">Customer contact communication type</td>
    <td class="tg-0pky">cellular, telephone</td>
  </tr>
  <tr>
    <td class="tg-0pky">month</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">Last contact month with the customer</td>
    <td class="tg-0pky">jan, feb, mar, ..., nov, dec</td>
  </tr>
  <tr>
    <td class="tg-0pky">day_of_week</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">Last contact day of the week with the customer <br>(note that the bank is only open during weekdays)</td>
    <td class="tg-0pky">mon, tue, wed, thu, fri</td>
  </tr>
  <tr>
    <td class="tg-0pky">duration</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Duration of the last contact with customer in seconds</td>
    <td class="tg-0pky">20, 30, 60</td>
  </tr>
  <tr>
    <td class="tg-0pky">campaign</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">The number of contacts performed during <br>the campaign for the customer</td>
    <td class="tg-0pky">3, 7, 15</td>
  </tr>
  <tr>
    <td class="tg-0pky">pdays</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Number of days passed since the customer was last contacted <br>(999 means customer was never contacted)</td>
    <td class="tg-0pky">20, 30, 999, etc.</td>
  </tr>
  <tr>
    <td class="tg-0pky">previous</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Number of contacts made with customer before this campaign</td>
    <td class="tg-0pky">20, 30, 1000, etc.</td>
  </tr>
  <tr>
    <td class="tg-0pky">poutcome</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">Outcome of the previous marketing campaign</td>
    <td class="tg-0pky">failure, nonexistent, success</td>
  </tr>
  <tr>
    <td class="tg-0pky">emp.var.rate</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Employment variation rate - quarterly indicator</td>
    <td class="tg-0pky">-1.2, 1.1, 2.3</td>
  </tr>
  <tr>
    <td class="tg-0pky">cons.price.idx</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Consumer price index - monthly indicator</td>
    <td class="tg-0pky">91.21, 93.33, 94.32</td>
  </tr>
  <tr>
    <td class="tg-0pky">cons.conf.idx</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Consumer confidence index - monthly indicator</td>
    <td class="tg-0pky">-50.8, -36.6, -26.9</td>
  </tr>
  <tr>
    <td class="tg-0pky">euribor3m</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Interest rate where a selection of European banks <br>lend one another funds in euros where <br>loans have a maturity of 3 months</td>
    <td class="tg-0pky">1.048, 4.857, 3.569</td>
  </tr>
  <tr>
    <td class="tg-0pky">nr.employed</td>
    <td class="tg-0pky">numeric</td>
    <td class="tg-0pky">Quarterly average of the total number of employed citizens</td>
    <td class="tg-0pky">4198, 5191, 5325</td>
  </tr>
  <tr>
    <td class="tg-0pky">y</td>
    <td class="tg-0pky">categorical</td>
    <td class="tg-0pky">Has the client subscribed to a term deposit? (the product)</td>
    <td class="tg-0pky">yes, no</td>
  </tr>
</table>


```{python}
ax = sns.distplot(bankingDF["age"])
ax.grid(False)
ax
```

```{python}
fig, ax = plt.subplots(figsize=(15,10))
job_totals = bankingDF["job"].value_counts()
job_totals = pd.DataFrame({"Job Type": job_totals.index,
                         "Total": job_totals.values})

ax = sns.barplot(x = "Total", y="Job Type", data=job_totals,
                 palette="Blues_d")
ax.grid(False)
ax
```

```{python}
# those who accept by job category
job_yes = (bankingDF[bankingDF["y"] == "yes"]
    .groupby('job')
    .agg({'y':'count'})
    .rename(columns = {'y':'Subscribed'})
    .reset_index())

# total number of people in each job category
# regardless of subscription status
all_job = (bankingDF
    .groupby('job')
    .agg({'y':'count'})
    .rename(columns = {'y':'Job Total'})
    .reset_index())

# calculate prevalence rates
job_yes['Job Total'] = all_job["Job Total"]
job_yes["Prevalence Rate"] = job_yes["Subscribed"]/job_yes["Job Total"]
job_yes = job_yes.sort_values("Prevalence Rate", ascending=False)


chart = sns.barplot(job_yes["job"], job_yes["Prevalence Rate"],
           palette=sns.cubehelix_palette(12, reverse=True))

chart.set_xticklabels(
    chart.get_xticklabels(), 
    rotation=45, 
    horizontalalignment='right',
    fontweight='light',
    fontsize='large'
)
chart
```

```{python}
j = job_yes.set_index("job")
j
```

```{r echo=F}
data.frame(py$j)
```

```{python}
bankingDF.describe()
```


